---
title: "Predictive Standard Deviation"
author: "Ziang Zhang"
date: "2024-06-27"
output: workflowr::wflow_html
header-includes:
  - \usepackage{amsthm}
  - \newtheorem{corollary}{Corollary}
editor_options:
  chunk_output_type: console
---

## **Introduction**

Recall that the M-GP is defined by:
\[L_af(x) = \sigma \xi(x),\]
where $L_a = D^2 - \alpha(x)D$ and $\alpha(x) = -1/a(x+c)$. 
The monotone function implied by $\alpha(x)$ is $g = D^{-1}[\exp(D^{-1}\alpha)]$.

The t-IWP2 can be defined by a similar SDE:
\[\tilde{L}_af(x) = \sigma \xi(x),\]
where $\tilde{L}_a = g'(x)^{-3/2}L_\alpha$.

We want to put prior on the standard deviation parameter $\sigma$ hence on the function $f$ in an interpretable way. 
One way to do this is through the notion of $h$-step predictive standard deviation (PSD) which is defined as:
\[\sigma_x(h) = \text{SD}[f(x+h)|f(u):u\leq x].\]
For process defined by time invariant differential operator, the PSD is invariant to the choice of $x$. 
For example, the h-step PSD for an IWP-2 can be shown as $\sigma\sqrt{h^3/3}$ ([Zhang et.al, 2024](https://www.tandfonline.com/doi/full/10.1080/10618600.2023.2289532)).
However for the M-GP and t-IWP2, since the differential operator involves $\alpha(x)$, the PSD is dependent on $x$.

## **PSD for M-GP and t-IWP2**

In this section, we study the behavior of PSD for M-GP and t-IWP2. 
Let $G_t(\tau)$ denotes the Green function of the operator $L_a$, the PSD $\sigma_x(h)$ can be written as:
\[\sigma^2_x(h) = \int_{x}^{x + h}G^2_{x+h}(t)dt.\]
Recall for the M-GP, the Green function is given by:
\[G_t(\tau) = \frac{1}{g'(\tau)}[g(t)-g(\tau)].\]
Using property of the Green function (see [textbook](https://people.uncw.edu/hermanr/mat463/odebook/) Equation 8.36), we can show that the Green function of t-IWP2 is given by:
\[\tilde{G}_t(\tau) = g'(\tau)^{1/2}[g(t)-g(\tau)].\]

### *PSD of M-GP has a nice limit*

As mentioned above, the PSD of M-GP generally depends on both the location $x$ and the step size $h$.
In certain cases, we do not have a well-informed choice of the best location $x$ to put the prior on $\sigma$. 
Fortunately, we can show that the PSD of M-GP has a nice limit as $x\to \infty$, which defines a (limiting) PSD that does not require specifying the location $x$.

> **Corollary 1 (Limiting PSD of M-GP):**
>
> Assume $\alpha(x) = -\frac{1}{a(x+c)}$ where $a \neq 0$. For the M-GP defined by $\alpha(x)$ \[\sigma^2(h) = \lim_{x \to \infty}\sigma_x^2(h) = \frac{\sigma^2 h^3}{3}.\]

The proof of corollary 1 is given in the appendix. 
The corollary shows that although the PSD of M-GP $\sigma_x(h)$ depends on the location $x$, the limiting PSD $\sigma(h) = \lim_{x \to \infty} \sigma_x(h)$ is **exactly the same as the PSD of IWP-2**, which does not depend on the location $x$.

### *Limiting behavior for PSD of t-IWP2*

It might be tempting to think that the PSD of t-IWP2 also converges to the PSD of IWP2 as $x\to \infty$. 
However, the following corollary shows that this is not the case.

> **Corollary 2 (Limiting PSD of t-IWP2):**
>
> Assume $\alpha(x) = -\frac{1}{a(x+c)}$ where $a \geq 0$. For the t-IWP2 defined by a monotone transformation $m(x)$ such that $m'(x)/m''(x) = \alpha(x)$, the limiting PSD is \[\sigma^2(h) = \lim_{x \to \infty}\sigma_x^2(h) = 0.\]

The proof of corollary 2 is also given in the appendix.
This corollary implies as the starting time increases, the PSD of t-IWP2 converges to zero, which means there is no uncertainty in the functional form of the $f$ as $x\to \infty$.

In other words, the seemingly naive approach of transforming the coordinate of an IWP2 through a monotone function actually puts a **very strong prior** on the unknown function $f$.
Basically it assumes there will be (almost) **no deviation from the base model** $\text{span}\{1,m(x)\}$ at the right end of the domain when $x$ is large.



## **PSD Computation**

In this section, we will show how to compute the PSD for M-GP.

The function `PSD_compute` computes the PSD for M-GP, at any arbitrary location $x$ and step size $h$.

Assume we know the value of $\sigma$, then $\sigma_x(h)$ can be computed as:

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(Matrix)
library(knitr)
source("code/01-state-space.R")
```

```{r}
sigma <- 5
sigma * PSD_compute(x = 0, h = 2, c = 1, alpha = 2)
```

Now, suppose we have an Exponential prior on $\sigma_x(h)$ parametrized as:
\[
\mathbb{P}(\sigma_x(h) > u) = \text{Prob}.
\]
We can use the function `prior_conversion_mgp` to convert the prior to an Exponential prior on $\sigma$.

For example, assume we have a prior $\mathbb{P}(\sigma_0(1) > 5) = 0.1$, then the corresponding prior on $\sigma$ is computed as:
```{r}
prior_psd <- list(u = 5, prob = 0.1)
prior_conversion_mgp(prior_psd, x = 0, h = 1, c = 1, alpha = 2)
```

In this case, it corresponds to an Exponential prior $\mathbb{P}(\sigma > 9.828196) = 0.1$.

## **Appendix**

### *Proof of Corollary 1*

To prove the corollary, we will utilize the following two lemmas.

> **Lemma 1:** Assume $h\geq\tau>0$, then for any $x$:
> \[\lim_{x\to \infty} \frac{g(x+h)-g(x+\tau)}{g'(x+\tau)} = h-\tau.\]

*Proof of Lemma 1:*
Since $g$ is a continuously differentiable, by mean value theorem we know 
\[g(x+h) - g(x+\tau)=g'(x+\nu) (h-\tau),\]
where $\nu \in (x+\tau, x+h)$.

Assume without the loss of generality that $a >0$ so $g$ is concave function that is monotonically increasing. 
Then we have \[g'(x+h) (h-\tau) \leq g'(x+\nu) (h-\tau) \leq g'(x+\tau) (h-\tau).\]

Since by definition $h = g'$, it is clear for the right hand side we have \[\lim_{x\to \infty} \frac{g'(x+\tau) (h-\tau)}{g'(x+\tau)} = h-\tau.\]

For the left hand side, note that:

\begin{equation}
\begin{aligned}
\lim_{x \to \infty}\frac{g'(x+h)}{g'(x+\tau)} &= \lim_{x \to \infty}\exp \bigg(\int_{x+\tau}^{x+h} \alpha(t) dt\bigg) \\
&= \lim_{x \to \infty}\exp \bigg(\int_{x+\tau}^{x+h} \frac{-1}{a(t+c)} dt\bigg) \\
&= \lim_{x \to \infty}\exp \bigg(\frac{1}{a} \ln \bigg(\frac{x+\tau+c}{x+h+c}\bigg)\bigg) \\
&=  \lim_{x \to \infty} \bigg(\frac{x+\tau+c}{x+h+c}\bigg)^{1/a} = 1.
\end{aligned}
\end{equation}

So by squeezing theorem, we have \[\lim_{x\to \infty} \frac{g(x+h)-g(x+\tau)}{g'(x+\tau)} = h-\tau.\] 
This lemma is hence proved. \qed

> **Lemma 2:** Assume $h>0$, then for any $x$:
> \[\lim_{x\to \infty} \int_{x}^{x+h} G^2_{x+h}(t)dt = \int_{0}^{h} \lim_{x\to \infty} G^2_{x+h}(x + t)dt.\]

*Proof of Lemma 2:*
The proof of this lemma is a simple application of the dominated convergence theorem.

First it is clear through a simple change of variable that \[\int_{x}^{x+h} G^2_{x+h}(t)dt = \int_{0}^{h} G^2_{x+h}(x + t)dt.\]

Again, assume without the loss of generality that $a >0$ so $g$ is concave.
We have
\begin{equation}
\begin{aligned}
\left|G_{x+h}(x + t)\right| &= \left|\frac{g(x+h) - g(x + t)}{g'(x+t)}\right| \\
&\leq \left|\frac{g'(x+t)(h-t)}{g'(x+t)}\right| \\
&= |h - t|.
\end{aligned}
\end{equation}

Since $|h - t|$ is integrable over $[0,h]$, this lemma follows from the dominated convergence theorem.

With the above two lemmas, we can show that:
\begin{equation}
\begin{aligned}
\lim_{x\to \infty} \int_{x}^{x+h} G^2_{x+h}(t)dt &= \int_{0}^{h} \lim_{x\to \infty} G^2_{x+h}(x + t)dt \\
&= \int_{0}^{h} \bigg\{\lim_{x\to \infty} \frac{1}{g'(x+t)}[g(x+h)-g(x+t)]\bigg\}^2dt \\
&= \int_{0}^{h} (h-t)^2dt \\
&= \frac{h^3}{3}.
\end{aligned}
\end{equation}

The proof of corollary 1 is hence completed.


### *Proof of Corollary 2*

To prove this corollary, we first note that the Green function of t-IWP2 is given by:
\[\tilde{G}_t(\tau) = g'(\tau)^{1/2}[g(t)-g(\tau)].\]

When $a > 0$, we know $|g(x+h) - g(x+\tau)| \leq g'(x+\tau)(h-\tau)$ if $h\geq \tau$.
So we have:
\begin{equation}
\begin{aligned}
g'(x + \tau)^{1/2}[g(x + h)-g(x + \tau)] &\leq g'(x + \tau)^{3/2}(h-\tau)\\
&= (h-\tau) \exp \bigg[ - \int_{0}^{x+\tau} \frac{1}{a(t+c)}dt \bigg]^{3/2} \\
&= (h-\tau) \exp \bigg[ - \frac{1}{a} \ln \bigg(\frac{x+\tau+c}{c}\bigg) \bigg]^{3/2} \\
&= (h-\tau) \bigg(\frac{c}{x+\tau+c}\bigg)^{\frac{3}{2a}}. \\
\end{aligned}
\end{equation}

Therefore, we know \[0\leq\lim_{x\to\infty} \tilde{G}^2_{x+h}(x+\tau) \leq (h-\tau)^2\lim_{x\to\infty}  \bigg(\frac{c}{x+\tau+c}\bigg)^{3/a} = 0.\]

Finally, since $\tilde{G}^2_{x+h}(x+\tau) \leq (h-\tau)^2$ which is integrable over $[0,h]$, we can apply the dominated convergence theorem.
Therefore \[\lim_{x\to\infty} \int_{x}^{x+h}\tilde{G}^2_{x+h}(t)dt = \lim_{x\to\infty}\int_{0}^{h}\tilde{G}^2_{x+h}(x + \tau)d\tau = \int_{0}^{h} \left[\lim_{x\to\infty} \tilde{G}^2_{x+h}(x+\tau) \right]d\tau = 0. \]

This completes the proof of corollary 2. 


