---
title: "Simulation 1"
author: "Ziang Zhang"
date: "2024-06-19"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r message=FALSE, warning=FALSE, echo=FALSE}
library(tidyverse)
library(Matrix)
library(knitr)
source("code/01-state-space.R")
source("code/02-FEM.R")
source("code/03-sampling.R")
c <- 1
B <- 1000
n <- 200
true_psd = 2
sd_noise = 2
```

## **Introduction**

In this simulation study, we study the robustness of the tIWP2 and the mGP priors.


## **Simulation A: True model is mGP**

The simulation A is designed as follows, for each of the $B = 1000$ replications:

- $f$ is simulated from a mGP process with $a = 2$, with a true PSD $\sigma_0(5) = 2$.
- For $x_i$ equally spaced between $[0,20]$, $n = 200$ observations are simulated from $y_i = f(x_i) + \epsilon_i$ with $\epsilon_i \sim N(0, 2)$. 
- The posterior $\pi(f|\boldsymbol{y})$ is obtained using both the tIWP2 and the mGP priors, only using data with $x \in [0,10]$. The prior for the smoothing parameter is chosen as an Exponential prior centered at the true PSD $\sigma_0(5)$.
- The pointwise coverage probabilities of the 50\%, 80\%, and 95\% credible interval for $f$, as well as the mean squared error and the mean relative error of the posterior mean of $f$ are all recorded for each replication.

```{r, echo=FALSE}
set.seed(123)
TMB::compile("code/fitGP_known_sd.cpp")
dyn.load(TMB::dynlib("code/fitGP_known_sd"))
PSD_tIWP2_compute <- function(h, sd = 1, x){
  -(sd^2) * 8/3*(((3*h + 8)*x + 4*x^2 + 3*h + 4)*sqrt(h + x + 1) - (h^2 + (5*h + 8)*x + 4*x^2 + 5*h + 4)*sqrt(x + 1))/(sqrt(h + x + 1)*sqrt(x + 1))
}
# A function to simulate the data once:
sim_data_once <- function(n = 199, sd_noise = 2, psd_fun = 2){
  x <- seq(0, 20, length.out = (n+1))
  true_psd_factor <- PSD_compute(h = 5, x = 0)
  sd_fun <- psd_fun/true_psd_factor
  f_vec <- mGP_sim(t = x, alpha = 2, c = c, initial_vec = c(0,5), sd = sd_fun)[-1,2]
  y <- f_vec + rnorm(n, 0, sd_noise)
  data <- data.frame(x = x[-1], y = y, f = f_vec)
  return(data)
}
# data_sim <- sim_data_once(n)
# plot(data_sim$x, data_sim$f, type = "l", col = "blue", lwd = 2, xlab = "x", ylab = "f(x)")
# points(data_sim$x, data_sim$y, col = "black")

# A function to fit the model once:
fit_mGP_once <- function(data_sim, data_train, u){
  # construct fixed design matrix, assume a = 2 and c = 0.5
  X <- as(cbind(1, sqrt(data_train$x + c)), "dgCMatrix")
  
  # construct random design matrix
  B <- (Diagonal(1, n = 2*nrow(data_sim)))
  # take 1,3,5,...
  B <- B[seq(1, 2*nrow(data_sim), by = 2),]
  # then take only the first n_train rows:
  B <- B[1:nrow(data_train),]
  
  # construct penalty matrix
  P <- mGP_joint_prec(t_vec = data_sim$x, a = 2, c = c)
  logPdet <- determinant(P)$modulus
  
  # fit the model
  tmbdat <- list(
    y = data_train$y,
    X = X,
    B = B,
    P = P,
    logPdet = as.numeric(logPdet),
    betaprec = 0.001,
    sig = sd_noise,
    u = u,
    alpha = 0.5
  )
  
  tmbparams <- list(
    W = c(rep(0, (ncol(X) + ncol(B)))),
    theta = 0
  )
  
  ff <- TMB::MakeADFun(
    data = tmbdat, 
    parameters = tmbparams, 
    DLL = "fitGP_known_sd",
    random = "W",
    silent = TRUE
    )
  
  ff$he <- function(w) numDeriv::jacobian(ff$gr, w)
  fit <- aghq::marginal_laplace_tmb(ff, k = 4, startingvalue = 0)
  return(fit)
}
fit_tIWP_once <- function(data_sim, data_train, u){
  m <- function(x) {2*sqrt(c)*(sqrt(x+c) - sqrt(c))}
  # construct fixed design matrix, assume a = 2 and c = 0.5
  X <- as(cbind(1, m(data_train$x)), "dgCMatrix")
  
  # construct random design matrix
  B <- (Diagonal(1, n = 2*nrow(data_sim)))
  # take 1,3,5,...
  B <- B[seq(1, 2*nrow(data_sim), by = 2),]
  # then take only the first n_train rows:
  B <- B[1:nrow(data_train),]
  
  # construct penalty matrix
  P <- IWP_joint_prec(t_vec = m(data_sim$x))
  logPdet <- determinant(P)$modulus
  
  # fit the model
  tmbdat <- list(
    y = data_train$y,
    X = X,
    B = B,
    P = P,
    logPdet = as.numeric(logPdet),
    betaprec = 0.001,
    sig = sd_noise,
    u = u,
    alpha = 0.5
  )
  
  tmbparams <- list(
    W = c(rep(0, (ncol(X) + ncol(B)))),
    theta = 0
  )
  
  ff <- TMB::MakeADFun(
    data = tmbdat, 
    parameters = tmbparams, 
    DLL = "fitGP_known_sd",
    random = "W",
    silent = TRUE
    )
  
  ff$he <- function(w) numDeriv::jacobian(ff$gr, w)
  fit <- aghq::marginal_laplace_tmb(ff, k = 4, startingvalue = 0)
  return(fit)
}
evaluate_model_once <- function(data_train, data_sim, model_fit, M = 3000, model = "mGP"){
  samps <- aghq::sample_marginal(quad = model_fit, M = M)
  # take 1,3,5..
  f_samps <- samps$samps[seq(1, nrow(data_sim)*2, by = 2),]
  beta_samps <- samps$samps[(nrow(data_sim)*2 + 1): (nrow(data_sim)*2 + 2),]
  if(model == "mGP"){
      X_refined <- as(cbind(1, sqrt(data_sim$x + c)), "dgCMatrix")
  }
  else{
    m <- function(x) {2*sqrt(c)*(sqrt(x+c) - sqrt(c))}
    X_refined <- as(cbind(1, m(data_sim$x)), "dgCMatrix")
  }

  f_samps_combined <- f_samps + X_refined %*% beta_samps
  
  ## interpolated:
  interval_50 <- apply(f_samps_combined, 1, function(x) quantile(x, c(0.25, 0.75)))
  coverage_50_inter <- mean(data_sim$f[1:nrow(data_train)] >= interval_50[1,1:nrow(data_train)] & data_sim$f[1:nrow(data_train)] <= interval_50[2,1:nrow(data_train)])
  coverage_50_pred <- mean(data_sim$f[-(1:nrow(data_train))] >= interval_50[1,-(1:nrow(data_train))] & data_sim$f[-(1:nrow(data_train))] <= interval_50[2,-(1:nrow(data_train))])
  
  interval_80 <- apply(f_samps_combined, 1, function(x) quantile(x, c(0.1, 0.9)))
  coverage_80_inter <- mean(data_sim$f[1:nrow(data_train)] >= interval_80[1,1:nrow(data_train)] & data_sim$f[1:nrow(data_train)] <= interval_80[2,1:nrow(data_train)])
  coverage_80_pred <- mean(data_sim$f[-(1:nrow(data_train))] >= interval_80[1,-(1:nrow(data_train))] & data_sim$f[-(1:nrow(data_train))] <= interval_80[2,-(1:nrow(data_train))])
  
  interval_95 <- apply(f_samps_combined, 1, function(x) quantile(x, c(0.025, 0.975)))
  coverage_95_inter <- mean(data_sim$f[1:nrow(data_train)] >= interval_95[1,1:nrow(data_train)] & data_sim$f[1:nrow(data_train)] <= interval_95[2,1:nrow(data_train)])
  coverage_95_pred <- mean(data_sim$f[-(1:nrow(data_train))] >= interval_95[1,-(1:nrow(data_train))] & data_sim$f[-(1:nrow(data_train))] <= interval_95[2,-(1:nrow(data_train))])
  # rmse of posterior mean
  rmse_inter <- sqrt(mean((rowMeans(f_samps_combined)[1:nrow(data_train)] - data_sim$f[1:nrow(data_train)])^2))
  rmse_pred <- sqrt(mean((rowMeans(f_samps_combined)[-(1:nrow(data_train))] - data_sim$f[-(1:nrow(data_train))])^2))
  
  # mean relative error of posterior mean
  rmle_inter <- mean(abs((rowMeans(f_samps_combined)[1:nrow(data_train)] - data_sim$f[1:nrow(data_train)])/data_sim$f[1:nrow(data_train)]))
  rmle_pred <- mean(abs((rowMeans(f_samps_combined)[-(1:nrow(data_train))] - data_sim$f[-(1:nrow(data_train))])/data_sim$f[-(1:nrow(data_train))]))
  # return the results as two dataframes
  return(
    list(
      data_inter = data.frame(coverage_50 = coverage_50_inter, coverage_80 = coverage_80_inter, coverage_95 = coverage_95_inter, rmse = rmse_inter, rmle = rmle_inter),
      data_pred = data.frame(coverage_50 = coverage_50_pred, coverage_80 = coverage_80_pred, coverage_95 = coverage_95_pred, rmse = rmse_pred, rmle = rmle_pred)
    )
  )

}

## A function to compare the result for B times, and aggregate the result into dataframes
replicate_comparison_B_times <- function(B, n, sd_noise = 2, psd_fun = 2, observed_max = 10, parallel = FALSE, u_mgp, u_tiwp2) {
  if (!parallel) {
    # (Sequential version remains the same)
    # Initialize empty data frames
    mGP_inter <- data.frame(coverage_50 = NULL, coverage_80 = NULL, coverage_95 = NULL, rmse = NULL, rmle = NULL)
    tIWP2_inter <- data.frame(coverage_50 = NULL, coverage_80 = NULL, coverage_95 = NULL, rmse = NULL, rmle = NULL)
    mGP_pred <- data.frame(coverage_50 = NULL, coverage_80 = NULL, coverage_95 = NULL, rmse = NULL, rmle = NULL)
    tIWP2_pred <- data.frame(coverage_50 = NULL, coverage_80 = NULL, coverage_95 = NULL, rmse = NULL, rmle = NULL)

    for (i in 1:B) {
      tryCatch({
        # Simulate data
        data_sim <- sim_data_once(n = n, sd_noise = sd_noise, psd_fun = psd_fun)
        data_train <- data_sim %>% filter(x <= observed_max)

        # Fit models
        mGP_fit <- fit_mGP_once(data_sim = data_sim, data_train = data_train, u = u_mgp)
        tiwp2_fit <- fit_tIWP_once(data_sim = data_sim, data_train = data_train, u = u_tiwp2)

        # Evaluate models
        mGP_result <- evaluate_model_once(data_train = data_train, data_sim = data_sim, model_fit = mGP_fit, model = "mGP")
        tIWP2_result <- evaluate_model_once(data_train = data_train, data_sim = data_sim, model_fit = tiwp2_fit, model = "tIWP2")

        # Combine results
        mGP_inter <- rbind(mGP_inter, mGP_result$data_inter)
        mGP_pred <- rbind(mGP_pred, mGP_result$data_pred)
        tIWP2_inter <- rbind(tIWP2_inter, tIWP2_result$data_inter)
        tIWP2_pred <- rbind(tIWP2_pred, tIWP2_result$data_pred)

        if (i %% 10 == 0) cat("Completed iteration", i, "out of", B, "\n")
      }, error = function(e) {
        cat("Error at iteration", i, ":", e$message, "\n")
      })
    }

    return(list(
      mGP_inter = mGP_inter,
      mGP_pred = mGP_pred,
      tIWP2_inter = tIWP2_inter,
      tIWP2_pred = tIWP2_pred
    ))
  } else {
    # Parallel version
    library(parallel)
    cl <- makeCluster(detectCores() - 12)
    on.exit(stopCluster(cl)) # Ensure the cluster is stopped even if an error occurs

    # Export necessary objects and functions
    clusterExport(cl, list("sim_data_once", "fit_mGP_once", "fit_tIWP_once", "evaluate_model_once", "n", "c", "psd_factor_mgp", "psd_factor_tiwp2", "true_psd", "sd_noise"))

    # Setup environment and load libraries on workers
    clusterEvalQ(cl, {
      library(tidyverse)
      library(Matrix)
      library(TMB)
      source("code/01-state-space.R")
      source("code/02-FEM.R")
      source("code/03-sampling.R")
      TMB::compile("code/fitGP_known_sd.cpp")
      dyn.load(TMB::dynlib("code/fitGP_known_sd"))
      TRUE  # Return TRUE to indicate successful setup
    })

    # Function to process each iteration in parallel
    process_iteration <- function(i) {
      tryCatch({
        # Simulate data
        data_sim <- sim_data_once(n = n, sd_noise = sd_noise, psd_fun = psd_fun)
        data_train <- data_sim %>% filter(x <= observed_max)

        # Fit models
        mGP_fit <- fit_mGP_once(data_sim = data_sim, data_train = data_train, u = u_mgp)
        tiwp2_fit <- fit_tIWP_once(data_sim = data_sim, data_train = data_train, u = u_tiwp2)

        # Evaluate models
        mGP_result <- evaluate_model_once(data_train = data_train, data_sim = data_sim, model_fit = mGP_fit, model = "mGP")
        tIWP2_result <- evaluate_model_once(data_train = data_train, data_sim = data_sim, model_fit = tiwp2_fit, model = "tIWP2")

        list(
          mGP_inter = mGP_result$data_inter,
          mGP_pred = mGP_result$data_pred,
          tIWP2_inter = tIWP2_result$data_inter,
          tIWP2_pred = tIWP2_result$data_pred
        )
      }, error = function(e) {
        cat("Error in iteration", i, ":", e$message, "\n")
        NULL
      })
    }

    # Run the iterations in parallel
    results <- parLapply(cl, 1:B, process_iteration)

    # Combine the results from all iterations
    mGP_inter <- do.call(rbind, lapply(results, `[[`, "mGP_inter"))
    mGP_pred <- do.call(rbind, lapply(results, `[[`, "mGP_pred"))
    tIWP2_inter <- do.call(rbind, lapply(results, `[[`, "tIWP2_inter"))
    tIWP2_pred <- do.call(rbind, lapply(results, `[[`, "tIWP2_pred"))

    return(list(
      mGP_inter = mGP_inter,
      mGP_pred = mGP_pred,
      tIWP2_inter = tIWP2_inter,
      tIWP2_pred = tIWP2_pred
    ))
  }
}
```

```{r eval=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
psd_factor_tiwp2 <- true_psd/PSD_tIWP2_compute(h = 5, x = 0)
psd_factor_mgp <- true_psd/PSD_compute(h = 5, x = 0)
all_result <- replicate_comparison_B_times(B = B, n = n, parallel = T, u_mgp = psd_factor_mgp, u_tiwp2 = psd_factor_tiwp2, sd_noise = 2, psd_fun = true_psd)
save(all_result, file = "output/sim1_all_resultA.rda")
```

```{r, echo=FALSE}
load("output/sim1_all_resultA.rda")
```

Visualize the RMSE result for interpolation and prediction:
```{r, echo=FALSE}
rmse_results <- bind_rows(
  all_result$mGP_inter %>%
    mutate(model = "mGP", type = "inter"),
  all_result$tIWP2_inter %>%
    mutate(model = "tIWP2", type = "inter"),
  all_result$mGP_pred %>%
    mutate(model = "mGP", type = "pred"),
  all_result$tIWP2_pred %>%
    mutate(model = "tIWP2", type = "pred")
) %>%
  select(rmse, model, type) %>%
  pivot_longer(cols = starts_with("rmse"),
               names_to = "rmse_level",
               values_to = "rmse_value") %>%
  mutate(rmse_level = gsub("rmse_", "", rmse_level))

rmse_results %>% ggplot(aes(x = type, y = (rmse_value), fill = model)) +
  geom_boxplot(outliers = F) +
  scale_fill_manual(values = c("mGP" = "skyblue", "tIWP2" = "salmon")) +
  labs(title = "Comparison of RMSE",
       x = "Setting",
       y = "RMSE",
       fill = "Model") +
  scale_y_log10() +
  theme_minimal()
```

Visualize the MLE result for interpolation and prediction:
```{r, echo=FALSE}
rmle_results <- bind_rows(
  all_result$mGP_inter %>%
    mutate(model = "mGP", type = "inter"),
  all_result$tIWP2_inter %>%
    mutate(model = "tIWP2", type = "inter"),
  all_result$mGP_pred %>%
    mutate(model = "mGP", type = "pred"),
  all_result$tIWP2_pred %>%
    mutate(model = "tIWP2", type = "pred")
) %>%
  select(rmle, model, type) %>%
  pivot_longer(cols = starts_with("rmle"),
               names_to = "rmle_level",
               values_to = "rmle_value") %>%
  mutate(rmle_level = gsub("rmle_", "", rmle_level))

rmle_results %>% ggplot(aes(x = type, y = rmle_value, fill = model)) +
  geom_boxplot(outliers = F) +
  scale_fill_manual(values = c("mGP" = "skyblue", "tIWP2" = "salmon")) +
  labs(title = "Comparison of mean relative error",
       x = "Setting",
       y = "relative error",
       fill = "Model") +
  theme_minimal()

```

A table of the summary statistics:
```{r, echo=FALSE}
# Aggregating mean and standard deviation for inter results
inter_summary <- bind_rows(
  all_result$mGP_inter %>%
    summarise(
      model = "mGP",
      mean_coverage_50 = mean(coverage_50),
      sd_coverage_50 = sqrt(mean(coverage_50)*(1-mean(coverage_50))/n()),
      mean_coverage_80 = mean(coverage_80),
      sd_coverage_80 = sqrt(mean(coverage_80)*(1-mean(coverage_80))/n()),
      mean_coverage_95 = mean(coverage_95),
      sd_coverage_95 = sqrt(mean(coverage_95)*(1-mean(coverage_95))/n()),
      mean_rmse = mean(rmse),
      sd_rmse = sd(rmse),
      mean_rmle = mean(rmle),
      sd_rmle = sd(rmle)
    ),
  all_result$tIWP2_inter %>%
    summarise(
      model = "tIWP2",
      mean_coverage_50 = mean(coverage_50),
      sd_coverage_50 = sqrt(mean(coverage_50)*(1-mean(coverage_50))/n()),
      mean_coverage_80 = mean(coverage_80),
      sd_coverage_80 = sqrt(mean(coverage_80)*(1-mean(coverage_80))/n()),
      mean_coverage_95 = mean(coverage_95),
      sd_coverage_95 = sqrt(mean(coverage_95)*(1-mean(coverage_95))/n()),
      mean_rmse = mean(rmse),
      sd_rmse = sd(rmse),
      mean_rmle = mean(rmle),
      sd_rmle = sd(rmle)
    )
) %>%
  mutate(case = "Inter")

# Aggregating mean and standard deviation for pred results
pred_summary <- bind_rows(
  all_result$mGP_pred %>%
    summarise(
      model = "mGP",
      mean_coverage_50 = mean(coverage_50),
      sd_coverage_50 = sqrt(mean(coverage_50)*(1-mean(coverage_50))/n()),
      mean_coverage_80 = mean(coverage_80),
      sd_coverage_80 = sqrt(mean(coverage_80)*(1-mean(coverage_80))/n()),
      mean_coverage_95 = mean(coverage_95),
      sd_coverage_95 = sqrt(mean(coverage_95)*(1-mean(coverage_95))/n()),
      mean_rmse = mean(rmse),
      sd_rmse = sd(rmse),
      mean_rmle = mean(rmle),
      sd_rmle = sd(rmle)
    ),
  all_result$tIWP2_pred %>%
    summarise(
      model = "tIWP2",
      mean_coverage_50 = mean(coverage_50),
      sd_coverage_50 = sqrt(mean(coverage_50)*(1-mean(coverage_50))/n()),
      mean_coverage_80 = mean(coverage_80),
      sd_coverage_80 = sqrt(mean(coverage_80)*(1-mean(coverage_80))/n()),
      mean_coverage_95 = mean(coverage_95),
      sd_coverage_95 = sqrt(mean(coverage_95)*(1-mean(coverage_95))/n()),
      mean_rmse = mean(rmse),
      sd_rmse = sd(rmse),
      mean_rmle = mean(rmle),
      sd_rmle = sd(rmle)
    )
) %>%
  mutate(case = "Pred")

# Combine mean and standard deviation into a single column
format_with_sd <- function(mean_val, sd_val) {
  paste0(round(mean_val, 3), " (", round(sd_val, 3), ")")
}

# Create summary table for the "Inter" scenario
inter_summary_table <- inter_summary %>%
  mutate(
    Coverage_50 = format_with_sd(mean_coverage_50, sd_coverage_50),
    Coverage_80 = format_with_sd(mean_coverage_80, sd_coverage_80),
    Coverage_95 = format_with_sd(mean_coverage_95, sd_coverage_95),
    RMSE = format_with_sd(mean_rmse, sd_rmse),
    MLE = format_with_sd(mean_rmle, sd_rmle)
  ) %>%
  select(Model = model, `Coverage 50%` = Coverage_50, `Coverage 80%` = Coverage_80, `Coverage 95%` = Coverage_95, RMSE, MLE)

# Create summary table for the "Pred" scenario
pred_summary_table <- pred_summary %>%
  mutate(
    Coverage_50 = format_with_sd(mean_coverage_50, sd_coverage_50),
    Coverage_80 = format_with_sd(mean_coverage_80, sd_coverage_80),
    Coverage_95 = format_with_sd(mean_coverage_95, sd_coverage_95),
    RMSE = format_with_sd(mean_rmse, sd_rmse),
    MLE = format_with_sd(mean_rmle, sd_rmle)
  ) %>%
  select(Model = model, `Coverage 50%` = Coverage_50, `Coverage 80%` = Coverage_80, `Coverage 95%` = Coverage_95, RMSE, MLE)

# Display the tables using kable
inter_summary_table %>%
  kable(format = "markdown", caption = "Summary for Inter Scenario")

pred_summary_table %>%
  kable(format = "markdown", caption = "Summary for Pred Scenario")
```

Visualize the coverage probabilities for interpolation and prediction:
```{r, echo=FALSE}
# Prepare the data for the Inter scenario
inter_coverage_plot_data <- inter_summary %>%
  pivot_longer(cols = starts_with("mean_coverage"),
               names_to = "Coverage_Level",
               values_to = "Coverage") %>%
  mutate(SD = case_when(
    Coverage_Level == "mean_coverage_50" ~ sd_coverage_50,
    Coverage_Level == "mean_coverage_80" ~ sd_coverage_80,
    Coverage_Level == "mean_coverage_95" ~ sd_coverage_95
  )) %>%
  mutate(
    lower = Coverage - 3 * SD,
    upper = Coverage + 3 * SD,
    Coverage_Level = recode(Coverage_Level, 
                            "mean_coverage_50" = "50%", 
                            "mean_coverage_80" = "80%", 
                            "mean_coverage_95" = "95%")
  )

ggplot(inter_coverage_plot_data, aes(x = Coverage_Level, y = Coverage, fill = model)) +
  geom_point(position = position_dodge(width = 0.5), size = 3, shape = 21, color = "black") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, position = position_dodge(width = 0.5)) +
  geom_hline(aes(yintercept = as.numeric(sub("%", "", Coverage_Level)) / 100), linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = c("mGP" = "skyblue", "tIWP2" = "salmon")) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(breaks = c(0.5, 0.8, 0.95)) +
  labs(title = "Interval Plot of Coverage Rates (Inter Scenario)",
       x = "Coverage Level",
       y = "Coverage Rate",
       fill = "Model") +
  theme_minimal()


# Prepare the data for the Pred scenario
pred_coverage_plot_data <- pred_summary %>%
  pivot_longer(cols = starts_with("mean_coverage"),
               names_to = "Coverage_Level",
               values_to = "Coverage") %>%
  mutate(SD = case_when(
    Coverage_Level == "mean_coverage_50" ~ sd_coverage_50,
    Coverage_Level == "mean_coverage_80" ~ sd_coverage_80,
    Coverage_Level == "mean_coverage_95" ~ sd_coverage_95
  )) %>%
  mutate(
    lower = Coverage - 3 * SD,
    upper = Coverage + 3 * SD,
    Coverage_Level = recode(Coverage_Level, 
                            "mean_coverage_50" = "50%", 
                            "mean_coverage_80" = "80%", 
                            "mean_coverage_95" = "95%")
  )

ggplot(pred_coverage_plot_data, aes(x = Coverage_Level, y = Coverage, fill = model)) +
  geom_point(position = position_dodge(width = 0.5), size = 3, shape = 21, color = "black") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, position = position_dodge(width = 0.5)) +
  geom_hline(aes(yintercept = as.numeric(sub("%", "", Coverage_Level)) / 100), linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = c("mGP" = "skyblue", "tIWP2" = "salmon")) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(breaks = c(0.5, 0.8, 0.95)) +
  labs(title = "Interval Plot of Coverage Rates (Pred Scenario)",
       x = "Coverage Level",
       y = "Coverage Rate",
       fill = "Model") +
  theme_minimal()
```






## **Simulation B: True model is tIWP2**

The simulation B is designed similarly as the simulation A, except now:

- $f$ is simulated from a tIWP2 process generated by $\sqrt{x+c}$, with a true PSD $\sigma_0(5) = 2$.

```{r, echo=FALSE}
set.seed(123)
# Update the function to simulate the data once:
sim_data_once_B <- function(n = 199, sd_noise = 2, psd_fun = 2){
  x <- seq(0, 20, length.out = (n+1))
  m <- function(x) {2*sqrt(c)*(sqrt(x+c) - sqrt(c))}
  m_deriv <- function(x) {sqrt(c)/sqrt(x+c)}
  x_trans <- m(x)
  true_psd_factor <- PSD_tIWP2_compute(h = 5, x = 0)
  sd_fun <- psd_fun/true_psd_factor
  f_vec <- sim_IWp_Var(t = x_trans, p = 2, sd = sd_fun, 
                      initial_vec = c(0,5/m_deriv(0)))[-1,2]
  y <- f_vec + rnorm(n, 0, sd_noise)
  data <- data.frame(x = x[-1], y = y, f = f_vec)
  return(data)
}
true_psd <- PSD_tIWP2_compute(h = 5, x = 0, sd = 1)
psd_factor_mGP <- true_psd/(PSD_compute(h = 5, x = 0, c = c, alpha = 2))
# data_sim <- sim_data_once_B(n)
# plot(data_sim$x, data_sim$f, type = "l", col = "blue", lwd = 2, xlab = "x", ylab = "f(x)")
# points(data_sim$x, data_sim$y, col = "black")
replicate_comparison_B_times_B <- function(B, n, sd_noise = 2, psd_fun = 2, observed_max = 10, parallel = FALSE, u_mgp, u_tiwp2) {
  if (!parallel) {
    # (Sequential version remains the same)
    # Initialize empty data frames
    mGP_inter <- data.frame(coverage_50 = NULL, coverage_80 = NULL, coverage_95 = NULL, rmse = NULL, rmle = NULL)
    tIWP2_inter <- data.frame(coverage_50 = NULL, coverage_80 = NULL, coverage_95 = NULL, rmse = NULL, rmle = NULL)
    mGP_pred <- data.frame(coverage_50 = NULL, coverage_80 = NULL, coverage_95 = NULL, rmse = NULL, rmle = NULL)
    tIWP2_pred <- data.frame(coverage_50 = NULL, coverage_80 = NULL, coverage_95 = NULL, rmse = NULL, rmle = NULL)

    for (i in 1:B) {
      tryCatch({
        # Simulate data
        data_sim <- sim_data_once_B(n = n, sd_noise = sd_noise, psd_fun = psd_fun)
        data_train <- data_sim %>% filter(x <= observed_max)

        # Fit models
        mGP_fit <- fit_mGP_once(data_sim = data_sim, data_train = data_train, u = u_mgp)
        tiwp2_fit <- fit_tIWP_once(data_sim = data_sim, data_train = data_train, u = u_tiwp2)

        # Evaluate models
        mGP_result <- evaluate_model_once(data_train = data_train, data_sim = data_sim, model_fit = mGP_fit, model = "mGP")
        tIWP2_result <- evaluate_model_once(data_train = data_train, data_sim = data_sim, model_fit = tiwp2_fit, model = "tIWP2")

        # Combine results
        mGP_inter <- rbind(mGP_inter, mGP_result$data_inter)
        mGP_pred <- rbind(mGP_pred, mGP_result$data_pred)
        tIWP2_inter <- rbind(tIWP2_inter, tIWP2_result$data_inter)
        tIWP2_pred <- rbind(tIWP2_pred, tIWP2_result$data_pred)

        if (i %% 10 == 0) cat("Completed iteration", i, "out of", B, "\n")
      }, error = function(e) {
        cat("Error at iteration", i, ":", e$message, "\n")
      })
    }

    return(list(
      mGP_inter = mGP_inter,
      mGP_pred = mGP_pred,
      tIWP2_inter = tIWP2_inter,
      tIWP2_pred = tIWP2_pred
    ))
  } else {
    # Parallel version
    library(parallel)
    cl <- makeCluster(detectCores() - 12)
    on.exit(stopCluster(cl)) # Ensure the cluster is stopped even if an error occurs

    # Export necessary objects and functions
    clusterExport(cl, list("sim_data_once_B", "fit_mGP_once", "fit_tIWP_once", "evaluate_model_once", "n", "c", "psd_factor_mgp", "psd_factor_tiwp2", "true_psd", "sd_noise", "PSD_tIWP2_compute"))

    # Setup environment and load libraries on workers
    clusterEvalQ(cl, {
      library(tidyverse)
      library(Matrix)
      library(TMB)
      source("code/01-state-space.R")
      source("code/02-FEM.R")
      source("code/03-sampling.R")
      TMB::compile("code/fitGP_known_sd.cpp")
      dyn.load(TMB::dynlib("code/fitGP_known_sd"))
      TRUE  # Return TRUE to indicate successful setup
    })

    # Function to process each iteration in parallel
    process_iteration <- function(i) {
      tryCatch({
        # Simulate data
        data_sim <- sim_data_once_B(n = n, sd_noise = sd_noise, psd_fun = psd_fun)
        data_train <- data_sim %>% filter(x <= observed_max)

        # Fit models
        mGP_fit <- fit_mGP_once(data_sim = data_sim, data_train = data_train, u = u_mgp)
        tiwp2_fit <- fit_tIWP_once(data_sim = data_sim, data_train = data_train, u = u_tiwp2)

        # Evaluate models
        mGP_result <- evaluate_model_once(data_train = data_train, data_sim = data_sim, model_fit = mGP_fit, model = "mGP")
        tIWP2_result <- evaluate_model_once(data_train = data_train, data_sim = data_sim, model_fit = tiwp2_fit, model = "tIWP2")

        list(
          mGP_inter = mGP_result$data_inter,
          mGP_pred = mGP_result$data_pred,
          tIWP2_inter = tIWP2_result$data_inter,
          tIWP2_pred = tIWP2_result$data_pred
        )
      }, error = function(e) {
        cat("Error in iteration", i, ":", e$message, "\n")
        NULL
      })
    }

    # Run the iterations in parallel
    results <- parLapply(cl, 1:B, process_iteration)

    # Combine the results from all iterations
    mGP_inter <- do.call(rbind, lapply(results, `[[`, "mGP_inter"))
    mGP_pred <- do.call(rbind, lapply(results, `[[`, "mGP_pred"))
    tIWP2_inter <- do.call(rbind, lapply(results, `[[`, "tIWP2_inter"))
    tIWP2_pred <- do.call(rbind, lapply(results, `[[`, "tIWP2_pred"))

    return(list(
      mGP_inter = mGP_inter,
      mGP_pred = mGP_pred,
      tIWP2_inter = tIWP2_inter,
      tIWP2_pred = tIWP2_pred
    ))
  }
}

```

```{r eval=FALSE, message=FALSE, warning=FALSE, echo=FALSE}
psd_factor_tiwp2 <- true_psd/PSD_tIWP2_compute(h = 5, x = 0)
psd_factor_mgp <- true_psd/PSD_compute(h = 5, x = 0)
all_result <- replicate_comparison_B_times_B(B = B, n = n, parallel = T, u_mgp = psd_factor_mgp, u_tiwp2 = psd_factor_tiwp2, sd_noise = sd_noise, psd_fun = true_psd)
save(all_result, file = "output/sim1_all_resultB.rda")
```

```{r, echo=FALSE}
load("output/sim1_all_resultB.rda")
```

Visualize the RMSE result for interpolation and prediction:
```{r, echo=FALSE}
rmse_results <- bind_rows(
  all_result$mGP_inter %>%
    mutate(model = "mGP", type = "inter"),
  all_result$tIWP2_inter %>%
    mutate(model = "tIWP2", type = "inter"),
  all_result$mGP_pred %>%
    mutate(model = "mGP", type = "pred"),
  all_result$tIWP2_pred %>%
    mutate(model = "tIWP2", type = "pred")
) %>%
  select(rmse, model, type) %>%
  pivot_longer(cols = starts_with("rmse"),
               names_to = "rmse_level",
               values_to = "rmse_value") %>%
  mutate(rmse_level = gsub("rmse_", "", rmse_level))

rmse_results %>% ggplot(aes(x = type, y = (rmse_value), fill = model)) +
  geom_boxplot(outliers = F) +
  scale_fill_manual(values = c("mGP" = "skyblue", "tIWP2" = "salmon")) +
  labs(title = "Comparison of RMSE",
       x = "Setting",
       y = "RMSE",
       fill = "Model") +
  scale_y_log10() +
  theme_minimal()
```

Visualize the MLE result for interpolation and prediction:
```{r, echo=FALSE}
rmle_results <- bind_rows(
  all_result$mGP_inter %>%
    mutate(model = "mGP", type = "inter"),
  all_result$tIWP2_inter %>%
    mutate(model = "tIWP2", type = "inter"),
  all_result$mGP_pred %>%
    mutate(model = "mGP", type = "pred"),
  all_result$tIWP2_pred %>%
    mutate(model = "tIWP2", type = "pred")
) %>%
  select(rmle, model, type) %>%
  pivot_longer(cols = starts_with("rmle"),
               names_to = "rmle_level",
               values_to = "rmle_value") %>%
  mutate(rmle_level = gsub("rmle_", "", rmle_level))

rmle_results %>% ggplot(aes(x = type, y = rmle_value, fill = model)) +
  geom_boxplot(outliers = F) +
  scale_fill_manual(values = c("mGP" = "skyblue", "tIWP2" = "salmon")) +
  labs(title = "Comparison of mean relative error",
       x = "Setting",
       y = "relative error",
       fill = "Model") +
  theme_minimal()

```

A table of the summary statistics:
```{r, echo=FALSE}
# Aggregating mean and standard deviation for inter results
inter_summary <- bind_rows(
  all_result$mGP_inter %>%
    summarise(
      model = "mGP",
      mean_coverage_50 = mean(coverage_50),
      sd_coverage_50 = sqrt(mean(coverage_50)*(1-mean(coverage_50))/n()),
      mean_coverage_80 = mean(coverage_80),
      sd_coverage_80 = sqrt(mean(coverage_80)*(1-mean(coverage_80))/n()),
      mean_coverage_95 = mean(coverage_95),
      sd_coverage_95 = sqrt(mean(coverage_95)*(1-mean(coverage_95))/n()),
      mean_rmse = mean(rmse),
      sd_rmse = sd(rmse),
      mean_rmle = mean(rmle),
      sd_rmle = sd(rmle)
    ),
  all_result$tIWP2_inter %>%
    summarise(
      model = "tIWP2",
      mean_coverage_50 = mean(coverage_50),
      sd_coverage_50 = sqrt(mean(coverage_50)*(1-mean(coverage_50))/n()),
      mean_coverage_80 = mean(coverage_80),
      sd_coverage_80 = sqrt(mean(coverage_80)*(1-mean(coverage_80))/n()),
      mean_coverage_95 = mean(coverage_95),
      sd_coverage_95 = sqrt(mean(coverage_95)*(1-mean(coverage_95))/n()),
      mean_rmse = mean(rmse),
      sd_rmse = sd(rmse),
      mean_rmle = mean(rmle),
      sd_rmle = sd(rmle)
    )
) %>%
  mutate(case = "Inter")

# Aggregating mean and standard deviation for pred results
pred_summary <- bind_rows(
  all_result$mGP_pred %>%
    summarise(
      model = "mGP",
      mean_coverage_50 = mean(coverage_50),
      sd_coverage_50 = sqrt(mean(coverage_50)*(1-mean(coverage_50))/n()),
      mean_coverage_80 = mean(coverage_80),
      sd_coverage_80 = sqrt(mean(coverage_80)*(1-mean(coverage_80))/n()),
      mean_coverage_95 = mean(coverage_95),
      sd_coverage_95 = sqrt(mean(coverage_95)*(1-mean(coverage_95))/n()),
      mean_rmse = mean(rmse),
      sd_rmse = sd(rmse),
      mean_rmle = mean(rmle),
      sd_rmle = sd(rmle)
    ),
  all_result$tIWP2_pred %>%
    summarise(
      model = "tIWP2",
      mean_coverage_50 = mean(coverage_50),
      sd_coverage_50 = sqrt(mean(coverage_50)*(1-mean(coverage_50))/n()),
      mean_coverage_80 = mean(coverage_80),
      sd_coverage_80 = sqrt(mean(coverage_80)*(1-mean(coverage_80))/n()),
      mean_coverage_95 = mean(coverage_95),
      sd_coverage_95 = sqrt(mean(coverage_95)*(1-mean(coverage_95))/n()),
      mean_rmse = mean(rmse),
      sd_rmse = sd(rmse),
      mean_rmle = mean(rmle),
      sd_rmle = sd(rmle)
    )
) %>%
  mutate(case = "Pred")

# Combine mean and standard deviation into a single column
format_with_sd <- function(mean_val, sd_val) {
  paste0(round(mean_val, 3), " (", round(sd_val, 3), ")")
}

# Create summary table for the "Inter" scenario
inter_summary_table <- inter_summary %>%
  mutate(
    Coverage_50 = format_with_sd(mean_coverage_50, sd_coverage_50),
    Coverage_80 = format_with_sd(mean_coverage_80, sd_coverage_80),
    Coverage_95 = format_with_sd(mean_coverage_95, sd_coverage_95),
    RMSE = format_with_sd(mean_rmse, sd_rmse),
    MLE = format_with_sd(mean_rmle, sd_rmle)
  ) %>%
  select(Model = model, `Coverage 50%` = Coverage_50, `Coverage 80%` = Coverage_80, `Coverage 95%` = Coverage_95, RMSE, MLE)

# Create summary table for the "Pred" scenario
pred_summary_table <- pred_summary %>%
  mutate(
    Coverage_50 = format_with_sd(mean_coverage_50, sd_coverage_50),
    Coverage_80 = format_with_sd(mean_coverage_80, sd_coverage_80),
    Coverage_95 = format_with_sd(mean_coverage_95, sd_coverage_95),
    RMSE = format_with_sd(mean_rmse, sd_rmse),
    MLE = format_with_sd(mean_rmle, sd_rmle)
  ) %>%
  select(Model = model, `Coverage 50%` = Coverage_50, `Coverage 80%` = Coverage_80, `Coverage 95%` = Coverage_95, RMSE, MLE)

# Display the tables using kable
inter_summary_table %>%
  kable(format = "markdown", caption = "Summary for Inter Scenario")

pred_summary_table %>%
  kable(format = "markdown", caption = "Summary for Pred Scenario")
```

Visualize the coverage probabilities for interpolation and prediction:
```{r, echo=FALSE}
# Prepare the data for the Inter scenario
inter_coverage_plot_data <- inter_summary %>%
  pivot_longer(cols = starts_with("mean_coverage"),
               names_to = "Coverage_Level",
               values_to = "Coverage") %>%
  mutate(SD = case_when(
    Coverage_Level == "mean_coverage_50" ~ sd_coverage_50,
    Coverage_Level == "mean_coverage_80" ~ sd_coverage_80,
    Coverage_Level == "mean_coverage_95" ~ sd_coverage_95
  )) %>%
  mutate(
    lower = Coverage - 3 * SD,
    upper = Coverage + 3 * SD,
    Coverage_Level = recode(Coverage_Level, 
                            "mean_coverage_50" = "50%", 
                            "mean_coverage_80" = "80%", 
                            "mean_coverage_95" = "95%")
  )

ggplot(inter_coverage_plot_data, aes(x = Coverage_Level, y = Coverage, fill = model)) +
  geom_point(position = position_dodge(width = 0.5), size = 3, shape = 21, color = "black") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, position = position_dodge(width = 0.5)) +
  geom_hline(aes(yintercept = as.numeric(sub("%", "", Coverage_Level)) / 100), linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = c("mGP" = "skyblue", "tIWP2" = "salmon")) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(breaks = c(0.5, 0.8, 0.95)) +
  labs(title = "Interval Plot of Coverage Rates (Inter Scenario)",
       x = "Coverage Level",
       y = "Coverage Rate",
       fill = "Model") +
  theme_minimal()


# Prepare the data for the Pred scenario
pred_coverage_plot_data <- pred_summary %>%
  pivot_longer(cols = starts_with("mean_coverage"),
               names_to = "Coverage_Level",
               values_to = "Coverage") %>%
  mutate(SD = case_when(
    Coverage_Level == "mean_coverage_50" ~ sd_coverage_50,
    Coverage_Level == "mean_coverage_80" ~ sd_coverage_80,
    Coverage_Level == "mean_coverage_95" ~ sd_coverage_95
  )) %>%
  mutate(
    lower = Coverage - 3 * SD,
    upper = Coverage + 3 * SD,
    Coverage_Level = recode(Coverage_Level, 
                            "mean_coverage_50" = "50%", 
                            "mean_coverage_80" = "80%", 
                            "mean_coverage_95" = "95%")
  )

ggplot(pred_coverage_plot_data, aes(x = Coverage_Level, y = Coverage, fill = model)) +
  geom_point(position = position_dodge(width = 0.5), size = 3, shape = 21, color = "black") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, position = position_dodge(width = 0.5)) +
  geom_hline(aes(yintercept = as.numeric(sub("%", "", Coverage_Level)) / 100), linetype = "dashed", color = "gray50") +
  scale_fill_manual(values = c("mGP" = "skyblue", "tIWP2" = "salmon")) +
  coord_cartesian(ylim = c(0, 1)) +
  scale_y_continuous(breaks = c(0.5, 0.8, 0.95)) +
  labs(title = "Interval Plot of Coverage Rates (Pred Scenario)",
       x = "Coverage Level",
       y = "Coverage Rate",
       fill = "Model") +
  theme_minimal()
```
















