---
title: "state_space_debug"
author: "Ziang Zhang"
date: "2024-06-24"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---


```{r}
library(tidyverse)
library(Matrix)
source("code/01-state-space.R")
source("code/02-FEM.R")
source("code/03-sampling.R")
B = 1000
c <- 1.1
```

## Check the variance

When $a = 1/2$:
```{r}
set.seed(123)
alpha = 1/2
samps0 <- mGP_sim(mesh_size = 1, max_t = 5, alpha = alpha, c = c)
result0 <- samps0[,1:2]
result0_deriv <- samps0[,c(1,3)]
for (i in 1:B) {
  samps0 <- mGP_sim(mesh_size = 1, max_t = 5, alpha = alpha, c = c)
  result0 <- cbind(result0, samps0[,2])
  result0_deriv <- cbind(result0_deriv, samps0[,3])
}
var(result0[nrow(result0),-1]) # 7.218279
var(result0_deriv[nrow(result0),-1]) # 1.162498
```
The theoretical variance computed for the function: (quite close)
```{r}
mspline_cov(s = 5, t = 5, c = c, alpha = alpha) # 7.233719
mspline_deriv_cov(s = 5, t = 5, c = c, alpha = alpha) # 1.219767
```
Compare with the variance computed from the FEM approach:
```{r}
xvec <- seq(0,5, by = 1)
samps <- sampling_from_FEM(x = xvec, a = alpha, c = c, k = 100, region = c(0,5), n = B, boundary = T, accuracy = 0.001)
### Sample paths for the function
plot(samps[,1]~xvec, type = 'l', col = 1, ylab = "samps", ylim = c(-10,10))
for (i in 1:ncol(samps)) {
  lines(samps[,i]~xvec, type = 'l', col = i)
}
var(samps[length(xvec),]) # 7.223936
```


When $a = 1$:
```{r}
set.seed(123)
alpha = 1
samps0 <- mGP_sim(mesh_size = 1, max_t = 5, alpha = alpha, c = c)
result0 <- samps0[,1:2]
result0_deriv <- samps0[,c(1,3)]
for (i in 1:B) {
  samps0 <- mGP_sim(mesh_size = 1, max_t = 5, alpha = alpha, c = c)
  result0 <- cbind(result0, samps0[,2])
  result0_deriv <- cbind(result0_deriv, samps0[,3])
}
var(result0[nrow(result0),-1]) # 14.98212
var(result0_deriv[nrow(result0),-1]) # 1.96319
```
The theoretical variance computed for the function: (quite close)
```{r}
mspline_cov(s = 5, t = 5, c = c, alpha = alpha) # 14.9063
mspline_deriv_cov(s = 5, t = 5, c = c, alpha = alpha) # 2.02141
```
Compare with the variance computed from the FEM approach:
```{r}
set.seed(123)
xvec <- seq(0,5, by = 1)
samps <- sampling_from_FEM(x = xvec, a = alpha, c = c, k = 100, region = c(0,5), n = B, boundary = T, accuracy = 0.001)
### Sample paths for the function
plot(samps[,1]~xvec, type = 'l', col = 1, ylab = "samps", ylim = c(-10,10))
for (i in 1:ncol(samps)) {
  lines(samps[,i]~xvec, type = 'l', col = i)
}
var(samps[length(xvec),]) # 14.18516
```

When $a = -1$:
```{r}
set.seed(123)
alpha = -1
samps0 <- mGP_sim(mesh_size = 1, max_t = 5, alpha = alpha, c = c)
result0 <- samps0[,1:2]
result0_deriv <- samps0[,c(1,3)]
for (i in 1:B) {
  samps0 <- mGP_sim(mesh_size = 1, max_t = 5, alpha = alpha, c = c)
  result0 <- cbind(result0, samps0[,2])
  result0_deriv <- cbind(result0_deriv, samps0[,3])
}
var(result0[nrow(result0),-1]) # 187.9503
var(result0_deriv[nrow(result0),-1]) # 28.45382
```
The theoretical variance computed for the function: 
```{r}
mspline_cov(s = 5, t = 5, c = c, alpha = alpha) # 183.7121
mspline_deriv_cov(s = 5, t = 5, c = c, alpha = alpha) # 27.72727
```
Compare with the variance computed from the FEM approach:
```{r}
set.seed(123)
xvec <- seq(0,5, by = 1)
samps <- sampling_from_FEM(x = xvec, a = alpha, c = c, k = 100, region = c(0,5), n = B, boundary = T, accuracy = 0.001)
### Sample paths for the function
plot(samps[,1]~xvec, type = 'l', col = 1, ylab = "samps", ylim = c(-10,10))
for (i in 1:ncol(samps)) {
  lines(samps[,i]~xvec, type = 'l', col = i)
}
var(samps[length(xvec),]) # 174.9346
```


When $a = -2$:
```{r}
set.seed(12345)
alpha = -2
samps0 <- mGP_sim(mesh_size = 1, max_t = 5, alpha = alpha, c = c)
result0 <- samps0[,1:2]
result0_deriv <- samps0[,c(1,3)]
for (i in 1:B) {
  samps0 <- mGP_sim(mesh_size = 1, max_t = 5, alpha = alpha, c = c)
  result0 <- cbind(result0, samps0[,2])
  result0_deriv <- cbind(result0_deriv, samps0[,3])
}
var(result0[nrow(result0),-1]) # 81.62238
var(result0_deriv[nrow(result0),-1]) # 10.33549
```
The theoretical variance computed for the function: 
```{r}
mspline_cov(s = 5, t = 5, c = c, alpha = alpha) # 82.02848
mspline_deriv_cov(s = 5, t = 5, c = c, alpha = alpha) # 10.44917
```
Compare with the variance computed from the FEM approach:
```{r}
set.seed(12345)
xvec <- seq(0,5, by = 1)
samps <- sampling_from_FEM(x = xvec, a = alpha, c = c, k = 100, region = c(0,5), n = B, boundary = T, accuracy = 0.001)
### Sample paths for the function
plot(samps[,1]~xvec, type = 'l', col = 1, ylab = "samps", ylim = c(-10,10))
for (i in 1:ncol(samps)) {
  lines(samps[,i]~xvec, type = 'l', col = i)
}
var(samps[length(xvec),]) # 82.31469
```


When $a = -1/2$:
```{r}
set.seed(123)
alpha = -1/2
samps0 <- mGP_sim(mesh_size = 1, max_t = 5, alpha = alpha, c = c)
result0 <- samps0[,1:2]
result0_deriv <- samps0[,c(1,3)]
for (i in 1:B) {
  samps0 <- mGP_sim(mesh_size = 1, max_t = 5, alpha = alpha, c = c)
  result0 <- cbind(result0, samps0[,2])
  result0_deriv <- cbind(result0_deriv, samps0[,3])
}
var(result0[nrow(result0),-1]) # 1379.876
var(result0_deriv[nrow(result0),-1]) # 354.2751
```
The theoretical variance computed for the function: (quite close)
```{r}
mspline_cov(s = 5, t = 5, c = c, alpha = alpha) # 1347.178
mspline_deriv_cov(s = 5, t = 5, c = c, alpha = alpha) # 344.7195
```
Compare with the variance computed from the FEM approach:
```{r}
xvec <- seq(0,5, by = 1)
samps <- sampling_from_FEM(x = xvec, a = alpha, c = c, k = 100, region = c(0,5), n = B, boundary = T, accuracy = 0.001)
### Sample paths for the function
plot(samps[,1]~xvec, type = 'l', col = 1, ylab = "samps", ylim = c(-10,10))
for (i in 1:ncol(samps)) {
  lines(samps[,i]~xvec, type = 'l', col = i)
}
var(samps[length(xvec),]) # 1391.181
```




## Sanity check using alpha = 2

```{r}
alpha = 2
samps0 <- mGP_sim(mesh_size = 0.1, max_t = 10, alpha = alpha, c = c, initial_vec = c(sqrt(c),1/(2*sqrt(c))), sd = 0)
plot(samps0[,2] ~ samps0[,1], type = 'l')
# plot sqrt(x + c) as a comparison
x = seq(0,10, by = 0.1)
lines(sqrt(x + c) ~ x, type = 'l', col = 2)
max(abs(samps0[,2] - sqrt(samps0[,1] + c))) # 1.332268e-15
```


## Sanity check using alpha = 1

```{r}
alpha = 1
samps0 <- mGP_sim(mesh_size = 0.1, max_t = 10, alpha = alpha, c = c, initial_vec = c(log(c),1/c), sd = 0)
plot(samps0[,2] ~ samps0[,1], type = 'l')
x = seq(0,10, by = 0.1)
lines(log(x + c) ~ x, type = 'l', col = 2)
max(abs(samps0[,2] - log(samps0[,1] + c))) # 4.440892e-15
```


## Sanity check using alpha = -1
```{r}
alpha = -1
samps0 <- mGP_sim(mesh_size = 0.1, max_t = 10, alpha = alpha, c = c, initial_vec = c(c^2,2*c), sd = 0)
plot(samps0[,2] ~ samps0[,1], type = 'l')
x = seq(0,10, by = 0.1)
lines(I((x+c)^2) ~ x, type = 'l', col = 2)
```

## Sanity check using alpha = -2
```{r}
alpha = -2
samps0 <- mGP_sim(mesh_size = 0.1, max_t = 10, alpha = alpha, c = c, initial_vec = c(c^(1.5),1.5*(c^(0.5))), sd = 0)
plot(samps0[,2] ~ samps0[,1], type = 'l')
x = seq(0,10, by = 0.1)
lines(I((x+c)^1.5) ~ x, type = 'l', col = 2)
```



